---
title: "qtlcharts - Karl Broman"
author: "StephensLab"
date: "June 24, 2015"
output: html_document
---

Karl Broman has recently developed a cool R package called **qtlcharts**. The installation instructions are as follows:

```{r indtallation, eval=FALSE, echo=TRUE}
install.packages(c("qtl","htmlwidgets","devtools"))
library(devtools)
install_github("kbroman/qtlcharts", build_vignettes=TRUE);

```

For getting a good idea about this package, some recommendations are 

- [qtlcharts webpage](http://kbroman.org/qtlcharts/)

- [qtlcharts paper](http://www.ncbi.nlm.nih.gov/pubmed/25527287)

We start by loading the package into R

```{r load-package, echo=TRUE, eval=TRUE}

library(qtlcharts)

```

I present a demo of how we applied this package to our use for the eQTLs analysis. The data is a toy example data. We use two datasets.

```{r datasets, echo=TRUE, eval=TRUE}

gene_data = read.table("../data/gene_expression_data.txt");
eqtl_data = read.table("../data/gene_eqtl_data.txt");

## check if the gene names match

tissue_names=as.vector(as.matrix(read.table("../data/tissue_id_eqtl_study.txt")));

```

### iplot (gene vs gene)

scatter plot between two genes based on tissue wide expression values (log scale).

```{r iplot, echo=TRUE, eval=TRUE}

scatter_plot <- iplot(as.numeric(log(gene_data+1)[2,]),as.numeric(log(gene_data+1)[3,]),rep(1,44), tissue_names)
scatter_plot

```

### iplotCurves (genes against tissues)

A plot of gene expression across tissues for all genes and a plot of eQTL effect sizes across tissues for all eQTLs.

```{r iplotCurves, echo=TRUE, eval=TRUE}

geneplot <- iplotCurves(log(gene_data+1),chartOpts=list(curves_xlab="Tissue",curves_ylab="log gene expression"))
htmlwidgets::saveWidget(geneplot, file="../plots/gene_expr_Curves.html",selfcontained=FALSE);

eqtlplot <- iplotCurves(eqtl_data,chartOpts=list(curves_xlab="Tissue",curves_ylab="eQTL strength effect"))
htmlwidgets::saveWidget(eqtlplot, file="../plots/eqtl_Curves.html",selfcontained=FALSE);

```

### iplotCorr (correlation structure)

construct a correlation heat map of the tissues based on gene expression data

```{r iplotCorr, echo=TRUE, eval=TRUE}

iplotCorr(gene_data,chartOpts=list(corcolors = c("white", "white","white", "crimson"),zlim=c(0,0.5,0.7,1),margin = list(left=150, top=40, right=3, bottom=70, inner=5)),reorder=TRUE)

```

### iBoxplot (box plot of the values)

construct a box plot of the effect sizes across different tissues for one eQTL.

```{r boxplot, echo=TRUE, eval=TRUE}

box_plot <- iboxplot(t(log(gene_data+1)),qu=c(0.25,0.05,0.01),chartOpts = list(qucolors=c("black","green","red","blue")),orderByMedian = FALSE)
htmlwidgets::saveWidget(box_plot, file="../plots/boxmap.html",selfcontained=FALSE);


```

### iplotMap

creates an interactive genetic map.

```{r iplotMap}
library(qtl)
data(hyper)
map <- pull.map(hyper)
genetic_map <- iplotMap(map)
htmlwidgets::saveWidget(genetic_map, file="../plots/genemap.html",selfcontained=FALSE);

```

### iHeatMap

construct a heat map of the tissues along with pixels linked to plots of horizontal and vertical slices. 

```{r iHeatMap, echo=TRUE, eval=TRUE}

cor_mat=cor(log(gene_data)+1);
x=1:44; y=1:44;

heatmap_plot <- iheatmap(cor_mat,x,y,chartOpts=list(corcolors = c("green", "white","crimson"),zlim=c(-1,0,1),margin = list(left=150, top=40, right=3, bottom=70, inner=5)))
htmlwidgets::saveWidget(heatmap_plot, file="../plots/heatmap.html",selfcontained=FALSE);

```




